Full Brain BOLD Signal Parameter Estimation \\ using Particle Filters

### 2010-08-25 doc/presentation/main.tex
outline of defense presentation done

\title{Full Brain Blood-Oxygen-Level-Dependent Signal \\
Parameter Estimation using Sequential Monte Carlo Methods}

@@

%\begin{frame}{BOLD Signal Properties}
%  \begin{itemize}
%    \item Exact variables and parameters are unknown and are
%        difficult to calculate.
%    \item Significant Amount of Lag between activation
%        and a measurable output - can be as much as 8 seconds.
%    \item Slow Temporal Resolution
%    \item Noise characterized by brownian motion, which clashes with low 
%        frequency elements.
%    \begin{center}
%    \includegraphics[width=10cm,height=4cm]{signal_fft.pdf}
%    \end{center}
%  \end{itemize}
%\end{frame}
%
%\begin{frame}{Preprocessing}
%  \begin{itemize}
%    \item Low Pass Filter (Gaussian Filter, not recommended)
%    \item Drift Removal (not always performed)
%    \begin{itemize}
%        \item High Pass Filter
%        \item Linear 
%        \item Quadratic
%        \item Wavelet 
%        \item Spline (Which I am using)
%    \end{itemize}
%  \end{itemize}
%  \begin{figure}
%    \includegraphics[width=10cm]{detrending_effect.png}
%    \caption{
%        \tiny
%        \cite{detrending}
%    }
%  \end{figure}
%\end{frame}
%
%\section{Statistical Parametric Mapping}
%\begin{frame}{Method}
%  \begin{figure}
%    \includegraphics[width=11cm]{glm_pipeline.png}
%    \caption{
%        \tiny
%        \cite{spm_pipeline}
%    }
%  \end{figure}
%\end{frame}
%
%\begin{frame}{Limitations}
%\begin{columns}
%  \column{2.5in}
%  \begin{itemize}
%    \item Linear, for a signal which is known to be nonlinear
%    \item Essentially the weighted sum of a set of "expected" responses.
%    \item Parametric
%    \begin{itemize}
%        \item Forced to make assumptions about underlying distributions
%        \item No time-scaling.
%    \end{itemize}
%  \end{itemize}
%  \column{3in}
%  \begin{figure}
%    \includegraphics[width=2.5in]{glm.png}
%    \caption{
%        \tiny
%        \cite{spm_pipeline}
%    }
%  \end{figure}
%\end{columns}
%\end{frame}
%
%\section{Nonlinear Regression}
%
%
%\section{Parameter Identification}
%\begin{frame}{Particle Filters}
%\begin{itemize}
%    \item Non-parametric, no assumptions are violated
%    \item Model based, fit parameters to input, constrained by physical variables
%    \item Fits a mixture PDF to the posterior of all parameters
%    \item Non-trivial computation cost
%    \item I use a Regularized Particle Filter
%    \begin{enumerate}
%        \item Regularized Re-sampling prevents particles from de-generating into a 
%            small number of unique particles
%        \item Allows distributions to move more freely
%    \end{enumerate}
%\end{itemize}
%\end{frame}
%
%\begin{frame}{Particle Filter}
%    \begin{columns}
%    \column{1.8in}
%    \begin{itemize}
%        \item $S_t = \{p_{0,t}, ... , p_{N,t}\}$, the set of particles
%        \item $w_{i,t}$, weight of particle $p_{i,t}$
%        \item $y_t$, measurement at time $t$, there is not a $y_t$ for every $t$.
%        \item $f(p_{i,t},y_t)$, weighting function 
%        \item $s(p_{i,t})$, step function 
%    \end{itemize}
%    \column{3.2in}
%    \begin{algorithmic}
%     \STATE Draw $S_0$ from prior distribution 
%     \FOR{$t = 0:t_{step}:t_{end}$}
%        \FOR{each $p_{i,t-1} \in S_{t-1}$}
%           \STATE $p_{i,t} = s(p_{i,t-1})$ 
%           \IF{There is a measurement at time $t$}
%               \FOR{every $p_{i,t}$}
%                    \STATE $w_{i,t} = w_{i,t-1}f(p_{i,t},y_t)$
%               \ENDFOR 
%               \STATE Resample if weights are unevenly distributed\\
%           \ENDIF
%        \ENDFOR
%     \ENDFOR
%%                    \item If a weights are not evenly distributed enough, then resample
%     \end{algorithmic}
%     \end{columns}
%\end{frame}
%
%\subsection{Single-Region}
%\begin{frame}{Single Timeseries Results}
%  \begin{figure}
%    \includegraphics[height=3in]{noise.png}
%  \end{figure}
%\end{frame}
%
%\begin{frame}{Single Timeseries Results, Measurement Convergence}
%  \begin{figure}
%    \includegraphics[height=3in]{bold.png}
%  \end{figure}
%\end{frame}
%
%\begin{frame}{Single Timeseries Results, State Convergence}
%  \begin{figure}
%    \includegraphics[height=3in]{state.png}
%  \end{figure}
%\end{frame}
%
%\begin{frame}{Factors Affecting Convergence}
%    \begin{enumerate}
%        \item Weighting function
%            \begin{itemize}
%                \item Needs to be continuous and defined for any input, should go to 0
%                \item Too wide a weighting function results in under-sensitivity, 
%                            slow or no convergence
%                \item Too thin a weighting function reduces robustness to noise
%            \end{itemize}
%        \item How often re-sampling is done, re-sampling should be minimized
%            \begin{itemize}
%                \item Stratified Resampling can result in truncated tails on posterior
%                \item Regularized Resampling can result in reduced robustness to noise 
%            \end{itemize}
%        \item Number of particles
%            \begin{itemize}
%                \item More particles give higher fidelity of posterior
%            \end{itemize}
%    \end{enumerate}
%\end{frame}
%
%\subsection{Multi-Region}
%\begin{frame}{Parameter Map Generation/Simulation}
%\begin{columns}
%  \column{2.5in}
%  \begin{enumerate}
%    \item Generate a parameter map, with a set of parameters for each voxel
%    \item Simulate every set of parameters, and use as input to possum
%    \item Perform preprocessing (de-trend and normalize)
%    \item Run particle filter on every grey matter voxel in image, generating a new
%            parameter map
%  \end{enumerate}
%  \column{3in}
%  \begin{figure}
%    \includegraphics[width=2.8in]{equation.png}
%    \caption{
%    }
%  \end{figure}
%\end{columns}
%\end{frame}
%
%\begin{frame}{Simulation Results, $\tau_0$}
%  \begin{figure}
%    \includegraphics[height=3in]{results.png}
%    \caption{Percent Difference from "Actual"}
%  \end{figure}
%\end{frame}
%
%\begin{frame}{Simulation Results, multiple}
%  \begin{figure}
%    \includegraphics[height=3in]{results2.png}
%    \caption{Percent Difference from "Actual"}
%  \end{figure}
%\end{frame}
%
%\section{Conclusion}
%
%% All of the following is optional and typically not needed. 
%\appendix
%\section<presentation>*{\appendixname}
%\subsection<presentation>*{For Further Reading}

>>>>

\title{Full Brain BOLD Signal Parameter Estimation \\
using Particle Filters}

@@

\section{Particle Filter}
\begin{frame}{Construction}
\begin{itemize}
    \item Initialize Particles with Prior ($\alpha$):\\
        $$\{(x^i_0,w^i) : x^i_0 \sim \alpha(X), w^i = 
                \frac{1}{N_p}, i \in \{1, 2, ... , N_p\} \}$$
        $$\alpha(X) \stackrel{\text{\tiny{k=0}}}{\approx} P(x_k) 
                = \sum_{i=0}^{N_p} w^i\delta(x_k - x^i_k ) dx$$
    \item Particle Weights at time $t_k$:\\
        $$w^i_k \propto \frac{P(x^i_{0:k} | y_{0:k})}{q(x^i_{0:k} | y_{0:k})}$$
      \note{$q$ is the density/location of the particle}
      \note{$p$ is the actual probability }
    \item Adding Measurements:\\
        $$w^i_k \propto & w^i_{k-1}P(y_k| x_k) $$
      \note{Because of the proportion, note that the weights do not
            have to integrate to 1, and can be extremely large or extremely small}
\end{itemize}
\end{frame}

\begin{frame}{Details and Example}
  \begin{columns}
    \begin{column}{.55\textwidth}
        \begin{itemize}
            \item Non-parametric
            \item Model based
            \item Online
            \item Provides Full Posterior Probability
            \item Non-trivial computation cost
            \item Regularized Particle Filter Used
        \end{itemize}
    \end{column}
    \begin{column}{.45\textwidth}
        \begin{figure}
          \includegraphics[width=\textwidth]{particle_filter}
        \end{figure}
    \end{column}
    \end{columns}
\end{frame}

%todo I should probably add labels on the axes
\section{Methods}
\begin{frame}{FMRI Noise}
\centering

Resting State Noise

\includegraphics[trim=3cm 0cm 3cm 0cm,width=.75\textwidth]{noise2_0009_22_38_23}

Resting State Noise Steps

\includegraphics[trim=3cm 0cm 3cm 0cm,width=.75\textwidth]{noise2_0009d_22_38_23}

\note{Tanabe 2002 \autoref{Tanabe2002} 10-15\% of voxels exhibit significant drift}
\note{Occurs in Cadavers.}
\end{frame}

\begin{frame}{Preprocessing}
  \begin{columns}
    \begin{column}{.7\textwidth}
        \begin{itemize}
            \item Drop Inital Volumes (9 Here)
            \item Realign Over Time
            \item Detrend
            \item Gaussian Smoothing (SPM Only)
            \begin{itemize}
                \item Imposes Gaussianity
                \item Increases SNR
                \item Reduces Bonferroni Correction Requirement
            \end{itemize}
        \end{itemize}
    \end{column}

    \begin{column}{.3\textwidth}
    %todo graphic of preprocessing
    \end{column}
  \end{columns}
\end{frame}

\section{Results}
\begin{frame}{Long Run Results}
%convergence
%correlation
\end{frame}

\begin{frame}{Short Run Results}
%results image
%table of estimated parameters
\end{frame}

\begin{frame}{POSSUM Results}
%mutual information map
\end{frame}

\begin{frame}{SPM vs. Mutual Information Map}

\end{frame}

\begin{frame}{Factors Affecting Convergence}
    \begin{itemize}
        \item Weighting function
            \begin{itemize}
                \item Continuous, Long Tailed, Zero-Mean 
                \begin{itemize}
                \item Too wide -> under-sensitivity, slow or no convergence
                \item Too thin -> reduces robustness to noise, particle deprivation
                \end{itemize}
            \end{itemize}
        \item Resampling 
            \begin{itemize}
                \item Stratified Resampling can result in truncated tails 
                        on posterior
                \item Regularized Resampling can result in over smoothing and 
                        slow convergence
            \end{itemize}
        \item Number of particles
            \begin{itemize}
                \item More particles give higher fidelity posterior
            \end{itemize}
    \end{itemize}
\end{frame}

\section{Conclusion}
\begin{frame}{Key Findings}

\end{frame}

\begin{frame}{Conclusion}
\end{frame}

%% All of the following is optional and typically not needed. 
%\appendix
%\section<presentation>*{\appendixname}
%\subsection<presentation>*{For Further Reading}
